---
title: "tidytuesday_stations"
author: "Mo"
date: "04/03/2022"
output: html_document
editor_options: 
  chunk_output_type: console
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	fig.height = 5,
	fig.width = 8,
	message = FALSE,
	warning = FALSE,
	cache = TRUE,
	cache.lazy = FALSE,
	dpi = 180
)
library(tidyverse)
library(silgelib)
library(tvthemes)

theme_set(theme_plex())
update_geom_defaults("rect", list(fill = "midnightblue", alpha = 0.8))
Sys.setlocale("LC_ALL","English")

setwd("C:/Data Science/2022/2022_03/freedom")
```

# Explore data
Our goal is 
```{r}
#tuesdata <- tidytuesdayR::tt_load('2022-02-22')
#freedom <- tuesdata$freedom
#write_csv(freedom,"C:/Data Science/2022/2022_03/freedom/freedom.csv")
freedom <- read_csv("freedom.csv")

freedom_df <- 
  freedom %>% 
  janitor::clean_names() %>%
  mutate(status = case_when(status == "F" ~ "free",
                            status == "NF" ~ "not free",
                            TRUE ~ "partially free")) %>%
  rename(civil_liberty = cl,
         political_right = pr,
         least_developed_country = is_ldc) %>%
  mutate(civil_liberty = case_when(
    civil_liberty == 1 ~ "cl_1",
    civil_liberty == 2 ~ "cl_2",
    civil_liberty == 3 ~ "cl_3",
    civil_liberty == 4 ~ "cl_4",
    civil_liberty == 5 ~ "cl_5",
    civil_liberty == 6 ~ "cl_6",
    TRUE ~ "cl_7"
  )) %>%
  mutate(political_right = case_when(
    political_right == 1 ~ "pl_1",
    political_right == 2 ~ "pl_2",
    political_right == 3 ~ "pl_3",
    political_right == 4 ~ "pl_4",
    TRUE ~ "pl_5"
  )) %>%
  mutate(least_developed_country = if_else(
    least_developed_country == 1, "least developed", "no least developed"
  )) %>% select(-region_code)

freedom_df
```

## civil liberty in general
```{r}
freedom_df %>% 
  count(year, civil_liberty, sort = T) %>%
  mutate(civil_liberty = factor(civil_liberty)) %>%
  ggplot(aes(year, n)) +
  geom_point(aes(col = civil_liberty), show.legend = F) +
  geom_line(aes(col = civil_liberty), show.legend = F) +
  facet_wrap(~civil_liberty, scales = "free_y", ncol = 2) +
  scale_color_avatar() +
  labs(y = "",
       title = "Civil liberty by year")
```

We notice a spike of cl_1 in mid 2001 till 2005 and slightly stabilizes
than a sharp plunge of cl_7 and the rest remain almost steady. 
## political right 
```{r}
freedom_df %>%
  count(year, political_right, sort = T) %>%
  mutate(political_right = factor(political_right)) %>%
  ggplot(aes(year, n)) +
  geom_point(aes(col = political_right), show.legend = F) +
  geom_line(aes(col = political_right), show.legend = F) +
  facet_wrap(~political_right, scales = "free_y", ncol = 2) +
  scale_color_avatar() +
  labs(y = "",
       title = "Political right by year")
```
The political rights oscillate during the years, probably  because of a lack of a major drive.

## least developed country
```{r}
freedom_df %>%
  select(civil_liberty, political_right, region_name,
         status, least_developed_country) %>%
  mutate(least_developed_country =
           fct_infreq(least_developed_country)) %>%
  mutate_if(is.numeric, as.character) %>%
  pivot_longer(civil_liberty:status, names_to = "feature",
               values_to = "value") %>%
  ggplot(aes(y = value, fill = least_developed_country)) +
  geom_bar(position = "fill") +
  facet_grid(rows = vars(feature), scales = "free_y", space = "free_y") +
  theme(legend.position = "top") +
  scale_fill_brewer(palette = "Dark2") +
  scale_x_continuous(expand = expansion(mult = c(0, .01)), 
                     labels = scales::percent) +
  labs(x = "", y = "", fill = "",
       title = "Least developed country word wide")
```
We can't extract a relevant insight unless we filter across region name.

## least developed country in Africa
```{r}
freedom_df %>%
  filter(region_name == "Africa") %>%
  select(civil_liberty, political_right, 
         status, least_developed_country) %>%
  mutate(least_developed_country = fct_infreq(least_developed_country)) %>%
  mutate_if(is.numeric, as.character) %>%
  pivot_longer(civil_liberty:status, names_to = "feature",
               values_to = "value") %>%
  ggplot(aes(y = value, fill = least_developed_country)) +
  geom_bar(position = "fill") +
  facet_grid(rows = vars(feature), scales = "free_y", space = "free_y") +
  theme(legend.position = "top") +
  scale_fill_brewer(palette = "Dark2") +
  scale_x_continuous(expand = expansion(mult = c(0, .01)), 
                     labels = scales::percent) +
  labs(x = "", y = "", fill = "",
       title = "Least developed country in Africa")
```


```{r}
freedom_df %>% #count(region_name)
  filter(region_name %in% c("Africa", "Asia")) %>%
  group_by(region_name) %>%
  count(year, political_right, sort = T) %>% 
  ggplot(aes(year, n)) +
  geom_point(aes(col = political_right)) +
  geom_line(aes(col = political_right)) +
  facet_grid(cols = vars(region_name),
             rows = vars(political_right),
             scales = "free") +
  scale_color_avatar() +
  labs(y = "",
       title = "Political right comparaison between Africa and Asia") +
  theme(legend.position = "none")
```
Political right in these two regions are similar which indicates that the regimes adapted are converging.

## least developed country in Europe
```{r}
freedom_df %>% #count(region_name)
  filter(region_name %in% c("Africa", "Asia")) %>%
  group_by(region_name) %>%
  count(year, political_right, sort = T) %>% 
  ggplot(aes(year, n)) +
  geom_point(aes(col = political_right)) +
  geom_line(aes(col = political_right)) +
  facet_grid(cols = vars(region_name),
             rows = vars(political_right),
             scales = "free") +
  scale_color_avatar() +
  labs(y = "",
       title = "Political right in Africa vs Asia") +
  theme(legend.position = "none")

```
Political right in these two regions are similar which indicates that the regimes adapted are converging.


## Political right score by region
```{r}
freedom_df %>%
  mutate(political_right = as.character(political_right)) %>%
  ggplot(aes(y = region_name, fill = political_right)) +
  geom_bar(position = "fill") +
  scale_fill_brewer(palette = "Dark2") +
  scale_x_continuous(expand = expansion(mult = c(0, .01)), 
                     labels = scales::percent) +
  labs(x = "", y = "", fill = "",
       title = "Political right score by region")
```

## Status by region
```{r}
freedom_df %>%
  ggplot(aes(y = region_name, fill = status)) +
  geom_bar(position = "fill") +
  scale_fill_brewer(palette = "Dark2") +
  scale_x_continuous(expand = expansion(mult = c(0, .01)), 
                     labels = scales::percent) +
  labs(x = "", y = "", fill = "",
       title = "Status by region")
```

# Build a model 
Let’s start our modeling by setting up our “data budget.” We’ll stratify by our outcome status.
```{r}
library(tidymodels)

set.seed(123)
freedom_split <- 
  freedom_df %>%
  mutate(year = factor(year)) %>%
  select(-region_name) %>%
  initial_split(strata = status)
freedom_train <- training(freedom_split)
freedom_test <- testing(freedom_split)

set.seed(234)
freedom_folds <- vfold_cv(freedom_train, strata = status)
freedom_folds
```

Next, let’s set up our feature engineering. We will need to transform character variable into dummy variables. Let’s also downsample since our dataset is imbalanced.
```{r}
library(themis)
library(textrecipes)

freedom_rec <-
  recipe(status ~ ., data = freedom_train) %>%
  update_role(country, new_role = "country") %>%
  step_dummy(all_nominal_predictors(), one_hot = T) %>%
  step_downsample(status)

juiced <- freedom_rec %>% prep() %>% juice()
```

# multinom_reg
Then, let’s create our model specification for a lasso model. We need to use a model specification that can handle multiclass data, in this case multinom_reg().
```{r}
multi_spec <-
  multinom_reg(penalty = tune(), mixture = 1) %>%
  set_mode("classification") %>%
  set_engine("glmnet")

multi_spec
```

Now it’s time to put the preprocessing and model together in a workflow().
```{r}
freedom_wf <- workflow(freedom_rec, multi_spec)
freedom_wf
```

Since the lasso regularization penalty is a hyperparameter of the model (we can’t find the best value from fitting the model a single time), let’s tune over a grid of possible penalty parameters.
```{r}
freedom_grid <- grid_regular(penalty(range = c(-5, 0)), 
                             levels = 20)

doParallel::registerDoParallel()
set.seed(2021)
freedom_rs <-
  tune_grid(
    freedom_wf,
    freedom_folds,
    grid = freedom_grid,
    control = control_grid(save_pred = T)
  )

freedom_rs 
```

We can look at the top results manually as well.
```{r}
show_best(freedom_rs)

show_best(freedom_rs, metric = "accuracy") %>%
  select(mean)

autoplot(freedom_rs)
```

We have pretty interesting result

Let’s use last_fit() to fit one final time to the training data and evaluate one final time on the testing data, with the numerically optimal result from freedom_rs
```{r}
freedom_last_fit <-
  freedom_wf %>%
  finalize_workflow(select_best(freedom_rs, "roc_auc")) %>%
  last_fit(freedom_split)

freedom_last_fit %>% collect_metrics()
```

How did this model perform on the testing data, that was not used in tuning/training?
```{r}
collect_predictions(freedom_last_fit) %>% 
  roc_auc(status, .pred_free:'.pred_partially free')
# 0.989
```

```{r}
collect_predictions(freedom_last_fit) %>%
  conf_mat(status, .pred_class) %>%
  autoplot()
```

We can also visualize this with an ROC curve.
```{r}
collect_predictions(freedom_last_fit) %>%
  roc_curve(status, .pred_free:'.pred_partially free') %>%
  ggplot(aes(1 - specificity, sensitivity, color = .level)) +
  geom_abline(lty = 2, color = "gray80", size = 1.5) +
  geom_path(alpha = 0.8, size = 1.2) +
  coord_equal() +
  labs(color = NULL)
```

Predicting free comes at the top, still we did a great job.

What features are most important for this multinom_reg model?
```{r}
library(vip)

freedom_fit <- extract_fit_parsnip(freedom_last_fit)
vip(freedom_fit, geom = "point", num_features = 10)
```

# Confidence interval

Let’s make one exploratory plot to see how the characteristics of the commercials change over time.
```{r}
freedom_df <- freedom %>%
  janitor::clean_names() %>%
  mutate(status = case_when(status == "F" ~ 1,
                            status == "PF" ~ 2,
                            TRUE ~ 3)) 
freedom_df %>%
  select(year, cl:is_ldc) %>%
  select(-region_name, -region_code) %>%
  pivot_longer(cl:is_ldc) %>%
  group_by(year, name) %>%
  summarise(prop = mean(value)) %>%
  ungroup() %>%
  ggplot(aes(year, prop, color = name)) +
  geom_line(size = 1.2, show.legend = FALSE) +
  facet_wrap(~name, scales = "free_y") +
  scale_y_continuous() +
  labs(x = NULL, y = "")
```

```{r}
simple_mod <- lm(year ~ cl + pr +
                  is_ldc + status,
                data = freedom_df)

summary(simple_mod)
```


```{r}
set.seed(123)
freedom_intervals <- 
  reg_intervals(year ~ cl + pr +
                  is_ldc + status,
                data = freedom_df,
                type = "percentile",
                keep_reps = TRUE
)

freedom_intervals
```


```{r}
freedom_intervals %>%
  mutate(term = fct_reorder(term, .estimate)
  ) %>%
  ggplot(aes(.estimate, term)) +
  geom_vline(xintercept = 0, size = 1.5, lty = 2, color = "gray80") +
  geom_errorbarh(aes(xmin = .lower, xmax = .upper),
                 size = 1.5, alpha = 0.5, color = "midnightblue"
  ) +
  geom_point(size = 3, color = "midnightblue") +
  labs(
    x = "Increase in year for each characteristic",
    y = NULL
  )

freedom_intervals %>%
  mutate(term = fct_reorder(term, .estimate)
  ) %>%
  unnest(.replicates) %>%
  ggplot(aes(estimate, fill = term)) +
  geom_vline(xintercept = 0, size = 1.5, lty = 2,
             color = "gray50") +
  geom_histogram(alpha = 0.8, show.legend = FALSE) +
  facet_wrap(vars(term))

```

We have evidence that freedom worldwide is witnessing
less civil liberties, increase in least developed countries appearance, liberties are sliding to partial liberties and a good-for-nothing increase of political rights.
Since civil rights are disappearing, we are witnessing the emergence of dictatorship rule.














